### 2020-06-27 chapters/my-work.tex
End of another hell day
Some Marco feedback regarding Validation

\subsubsection*{Multiple Test Correction} \label{sec:multi-test}
Note that we periodically apply this test (Equation \ref{eq:alert-test}) for each feature, effectively repeating the analysis. When considering several hypotheses, the problem of multiplicity arises: the more hypotheses are checked, the higher the probability of obtaining false positives. In other words, when repeating a test multiple times the multiple test or simultaneous statistical inference problem \cite{MultiTestProblem-Pubmed}
\cite{MultiTestProblem-Dickhaus2014}
\cite{MultiTestProblem-Miller1966}  \cite{multiple-test-correction-geoffrey} arises which states that the more inferences are made, the more likely erroneous inferences are to occur. In the literature, many multiple test correction methods are proposed, but we opted to use the Holm-Bonferroni multiple test correction due to its simplicity and ease of computation and leave as future work the testing of other correction methods. In statistics, Family-Wise Error Rate (FWER) \cite{MultivariateMT} \cite{MultitestTamhane2018AdvancesIP} is the probability of making one or more false positives when performing multiple hypotheses tests, \textit{i.e.}, consider one hypothesis true when it is not. The Holm-Bonferroni method \cite{HolmBonferroni} is one of many approaches for controlling the Family-Wise Error Rate \textit{(FWER)} by adjusting the rejection criteria for each hypothesis.

@@

In \cite{SAMM}, the authors prove that the minimum number of samples $S$ needed to satisfy these probability requirements is:
\begin{equation}
    \label{eq:optimal-n-samples}
    \textit{S} = \frac{\log\gamma}{log(1 - \frac{\alpha}{n_{features}})}
\end{equation}

>>>>

\subsubsection*{Multiple Test Correction} \label{sec:multi-test}
Note that we periodically apply this test (Equation \ref{eq:alert-test}) for each feature, effectively repeating the analysis. When considering several hypotheses, the problem of multiplicity arises: the more hypotheses are checked, the higher the probability of obtaining false positives. In other words, when repeating a test multiple times the multiple test or simultaneous statistical inference problem \cite{MultiTestProblem-Pubmed}
\cite{MultiTestProblem-Dickhaus2014}
\cite{MultiTestProblem-Miller1966}  \cite{multiple-test-correction-geoffrey} arises which states that the more inferences are made, the more likely erroneous inferences are to occur. In the literature, many multiple test correction methods are proposed, but we opted to use the Holm-Bonferroni multiple test correction because it is a simple and easy to compute correction that can reduce the amount of false negatives while having as target to control the amount of false positives. Furthermore, it does not require to assume any indepedence between the tested hypothesis (as other methods do) which would almost surely be a wrong assumption for realistic datasets. Therefore we leave as future work the testing of other correction methods. In statistics, Family-Wise Error Rate (FWER) \cite{MultivariateMT} \cite{MultitestTamhane2018AdvancesIP} is the probability of having one or more false positives when performing multiple hypotheses tests, \textit{i.e.}, considering one hypothesis true when it is not. The Holm-Bonferroni method \cite{HolmBonferroni} is one of many approaches for controlling the Family-Wise Error Rate \textit{(FWER)} by adjusting the rejection criteria for each hypothesis.

@@

In \cite{SAMM}, the authors prove that the minimum number of samples $S$ needed to satisfy these probability requirements is:
\begin{equation}
    \label{eq:optimal-n-samples}
    \textit{S} = \ceil[\bigg]{\frac{\log\gamma}{log(1 - \frac{\alpha}{n_{features}})}}
\end{equation}

### 2020-06-27 chapters/validation.tex
Some Marco feedback regarding Validation

\subsubsection{Varying sample sizes}
We began with single-feature scenarios and by experimenting with different tuple-based half-lives $n_{1/2}$ which in turn controls the size of the samples we make. In Section \ref{sec:ema-hist}, we mentioned we discarded events whose contribution was less than 6\%. We set a half-life $n_{1/2}$ value and then multiply it by a constant factor of 4 to get the size of the samples to make (a larger sample would contain events whose contribution was equal or less than 6\%).

The reference dataset R1 used in this first experiment contained a single feature that followed a normal distribution with mean $\mu=10$ and standard deviation $\sigma=2$ for the 5 million rows of the dataset. Figure \ref{fig:timeseries-r1} shows the time-series of the single feature in dataset R1.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.6]{figures/01-reference.png}
      \caption[]{Time-series for dataset R1}
      \label{fig:timeseries-r1}
    \end{center}
\end{figure}
The dataset used to process as a data stream or the target dataset also had 5 million events with a single feature. For the first half, \textit{i.e.}, for the first 2.5 million events, the feature followed a normal distribution with mean $\mu=10$ and standard deviation $\sigma=2$, like the reference dataset. For the other half, we changed the generating distribution to be a continuous uniform one with lower bound $a=100$ and upper bound $b=200$. Figure \ref{fig:timeseries-t1} shows the time-series of the single feature in the target dataset T1 where we see the abrupt change in feature values.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.6]{figures/01-target.png}
      \caption[]{Time-series for dataset T1}
      \label{fig:timeseries-t1}
    \end{center}
\end{figure}

We conducted three different experiments varying the tuple-based half-life $n_{1/2}$, using the reference dataset R1 and the target dataset T1. These experiments, corresponding varying half-lifes and execution times are shown in Table \ref{tbl:tests-half-life}. Notice that $Sample Size$ corresponds to the size of each sample made in the batch phase and will be four times the half-life. We used 100 bins (plus the two special infinity bins) for the reference histogram and made 1000 samples.
\begin{table}[!htb]
    \begin{center}
        \begin{tabular}{|c|c|c|c|c|}
        \hline
        \multicolumn{1}{|l|}{\textbf{Experiment}} & \multicolumn{1}{l|}{\textbf{Half-life}} & \multicolumn{1}{l|}{\textbf{Sample Size}} & \multicolumn{1}{l|}{\textbf{Executors}} & \multicolumn{1}{l|}{\textbf{Batch execution time (seconds)}} \\ \hline
        01                                        & 625                                     & 2500                                      & 300                                     & 152                                                          \\ \hline
        02                                        & 62,500                                  & 250,000                                   & 300                                     & 2,822                                                        \\ \hline
        03                                        & 250,000                                 & 1,000,000                                 & 300                                     & 10,082                                                       \\ \hline
        \end{tabular}
    \end{center}
    \caption{Experiments with varying half-life}
    \label{tbl:tests-half-life}
\end{table}
7 6 8
Table \ref{tbl:tests-half-life} shows the execution time for the batch phase in seconds. All experiments used 300 Spark executor instances. The batch phase comprises more than the sampling process but the remaining tasks take the same time across experiment. From the Table we conclude that the larger the half-life, the larger the sample size. Larger samples imply longer processing times when computing the EMA-like histogram for each sample. With no surprise, we see that the larger the sample the longer it takes to finish the batch phase.

From a functional point of view, were we able to detect the abrupt change introduced in the target dataset T1? Recall that from the 2.5 millionth event forward we change the distribution from a gaussian to an uniform one (Figure \ref{fig:timeseries-t1}). Figure \ref{fig:JSD-signal-01} shows the distance (JSD) values computed during streaming at each event of the 5 million events from dataset T1, using the outputs from the batch phase of experiment 01. We clearly see a change in JSD values, and we move from values close to 0 to values close to 1 (or even 1). Figure \ref{fig:JSD-signal-zoom-01} shows the same signal but for the first half of the streaming period (essentially a zoom-in). We see our JSD values before the change were very low (magnitudes of $10^{-2}$).
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.6]{figures/stream-analysis-viz-625.png}
      \caption[]{Distance (JSD) signal and threshold for experiment 01}
      \label{fig:JSD-signal-01}
    \end{center}
\end{figure}
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.6]{figures/stream-analysis-viz-zoom-625.png}
      \caption[]{Zoomed in distance (JSD) signal and threshold for experiment 01}
      \label{fig:JSD-signal-zoom-01}
    \end{center}
\end{figure}
However, the change in JSD value is very abrupt as well. Almost instantly we go from close to 0 values to 1. We also had some false-positives, \textit{i.e.}, some JSD values above the threshold. This indicates to us that the window used is too small and sensible to small changes.

@@

\subsubsection{Multiple distribution changes, each contained within the previous one}
In this experiment, we wanted to see how our signal behaved when there were multiple change points, where each distribution was contained within the previous one. Figure \ref{fig:timeseries-r2} shows the time-series for the reference dataset R2 used in this experiment. It contained 5 million events with a single feature following a gaussian distribution with mean $\mu=100$ and standard deviation $\sigma=50$.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.6]{figures/timeseries-r2.png}
      \caption[]{Time-series for dataset R2}
      \label{fig:timeseries-r2}
    \end{center}
\end{figure}
The time-series for target dataset T2 is shown in Figure \ref{fig:timeseries-t2}. Notice that we change distributions at 2, 3 and 4 millionth events. Notice also that each subsequent distribution produces values whose domain is a subset of the previous distribution. 
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.6]{figures/timeseries-t2.png}
      \caption[]{Time-series for dataset T2}
      \label{fig:timeseries-t2}
    \end{center}
\end{figure}

@@

\subsubsection{Multiple distribution changes, each a superset of the previous one}

In this experiment, we wanted to do the reverse of the previous one. In this one, there are multiple change points but each new distribution is a superset of the previous one. Figure \ref{fig:timeseries-r3} shows the reference period used, with 5 million events and one feature following a gaussian distribution of mean $\mu=100$ and standard deviation $\sigma=10$.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.6]{figures/timeseries-r3.png}
      \caption[]{Time-series for dataset R3}
      \label{fig:timeseries-r3}
    \end{center}
\end{figure}
Figure \ref{fig:timeseries-t3} shows the target period used, also with 5 million events and one feature. Like in the previous experiment, we change distributions at the 2, 3 and 4 millionth event mark. On the other hand, unlike the previous experiment, we now make subsequent distributions cover the previous ones. We change the reference gaussian used with mean $\mu=100$ and standard deviation $\sigma=10$ at the 2, 3 and 4 millionth mark, to use standard deviations $\sigma=30$, $\sigma=40$ and $\sigma=50$, respectively.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.6]{figures/timeseries-t3.png}
      \caption[]{Time-series for dataset T3}
      \label{fig:timeseries-t3}
    \end{center}
\end{figure}

@@

\newpage

@@

Figure \ref{fig:JSD-signal-test04} shows the computed JSD signal for the target period.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.6]{figures/stream-analysis-viz-test04.png}
      \caption[]{Distance (JSD) signal and threshold}
      \label{fig:JSD-signal-test04}
    \end{center}
\end{figure}
As expected, our JSD signal increases at the 2 millionth and 3 millionth event marks, because the distribution parameters change. At the 4 millionth event, we use the original reference distribution and notice that the JSD signal returns to values below the threshold, considering it a normal state, as we intended. Figure \ref{fig:JSD-signal-zoom-test04} is a zoom in on the first 2 millionth events, before any change occurs, and we see no false positives.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.6]{figures/stream-analysis-viz-zoom-test04.png}
      \caption[]{Zoomed in distance (JSD) signal and threshold}
      \label{fig:JSD-signal-zoom-test04}
    \end{center}
\end{figure}

@@

The first alert we obtain is:
\[1001000, [(0.0019999568566211323, x2)]\]
The previous alert tells us that at event 1,001,000 the only diverging feature in the system is feature \textit{x2} and what the p-value for the feature value at that time is. This is consistent with our target period induced changes. We introduced a change in distribution at the 1 millionth event for feature x2. We caught this change 1000 events later because of our test frequency, which is adjustable to obtain results as recent as we need them to be.

>>>>

\subsubsection{Varying sample sizes}
We began with single-feature scenarios and by experimenting with different tuple-based half-lives $n_{1/2}$. The half-life $n_{1/2}$ parameter in turn controls the size of the samples we make in the batch phase and the target sliding window size of the streaming phase.

In Section \ref{sec:ema-hist}, we mentioned that we choose to discard events whose contribution is less than 6\%. We set a half-life $n_{1/2}$ value and then multiply it by a constant factor of 4 to get the size of the samples to make (a larger sample would contain events whose contribution was equal or less than 6\%).

All the synthetic datasets used throughout this Section are a time-series where at each time-step we sample uniformly from a fixed distribution, \textit{i.e.} there are no time correlations in the used datasets. The reference dataset R1 used in this first experiment contained a single feature that followed a normal distribution with mean $\mu=10$ and standard deviation $\sigma=2$ for the 5 million rows of the dataset. Figure \ref{fig:timeseries-r1} shows the time-series of the single feature in dataset R1.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.6]{figures/01-reference.png}
      \caption[]{Time-series for dataset R1}
      \label{fig:timeseries-r1}
    \end{center}
\end{figure}
The dataset processed in the streaming phase (\textit{i.e.} the target dataset) was generated with a size of 5 million events and a single feature. For the first half, \textit{i.e.}, for the first 2.5 million events, the feature followed a normal distribution with mean $\mu=10$ and standard deviation $\sigma=2$, like the reference dataset. For the other half, we changed the generating distribution to be a continuous uniform one with lower bound $a=100$ and upper bound $b=200$. Figure \ref{fig:timeseries-t1} shows the time-series of the single feature in the target dataset T1 where we see the abrupt change in feature values.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.6]{figures/01-target.png}
      \caption[]{Time-series for dataset T1}
      \label{fig:timeseries-t1}
    \end{center}
\end{figure}

We conducted three different experiments varying the tuple-based half-life $n_{1/2}$, using the reference dataset R1 and the target dataset T1. The experiment identifiers, the corresponding half-lives and the execution times are shown in Table \ref{tbl:tests-half-life}. Notice that $Sample Size$ corresponds to the size of each sample made in the batch phase and so it is four times the half-life. We used 100 bins (plus the two special infinity bins) for the reference histogram and made 1000 samples.
\begin{table}[!htb]
    \begin{center}
        \begin{tabular}{|c|c|c|c|c|}
        \hline
        \multicolumn{1}{|l|}{\textbf{Experiment}} & \multicolumn{1}{l|}{\textbf{Half-life}} & \multicolumn{1}{l|}{\textbf{Sample Size}} & \multicolumn{1}{l|}{\textbf{Executors}} & \multicolumn{1}{l|}{\textbf{Batch execution time (seconds)}} \\ \hline
        01                                        & 625                                     & 2500                                      & 300                                     & 152                                                          \\ \hline
        02                                        & 62,500                                  & 250,000                                   & 300                                     & 2,822                                                        \\ \hline
        03                                        & 250,000                                 & 1,000,000                                 & 300                                     & 10,082                                                       \\ \hline
        \end{tabular}
    \end{center}
    \caption{Experiments with varying half-life}
    \label{tbl:tests-half-life}
\end{table}
Table \ref{tbl:tests-half-life} shows the execution time for the batch phase in seconds. All experiments used 300 Spark executor instances. From the Table we conclude that the larger the half-life, the larger the sample size. Larger samples imply longer processing times when computing the EMA-like histogram for each sample. With no surprise, we see that the larger the sample the longer it takes to finish the batch phase.

From a functional point of view, were we able to detect the abrupt change introduced in the target dataset T1? Recall that from the 2.5 millionth event forward we change the distribution from a gaussian to an uniform one (Figure \ref{fig:timeseries-t1}). Figure \ref{fig:JSD-signal-01} shows the distance (JSD) values computed during streaming at each event of the 5 million events from dataset T1, using the outputs from the batch phase of experiment 01. We clearly see a change in JSD values, and we move from values close to 0 to values close to 1 (or even 1). Figure \ref{fig:JSD-signal-zoom-01} shows the same signal but for the first half of the streaming period (essentially a zoom-in). We see our JSD values before the change were very low (magnitudes of $10^{-2}$).
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.6]{figures/stream-analysis-viz-625.png}
      \caption[]{Distance (JSD) signal and threshold for experiment 01}
      \label{fig:JSD-signal-01}
    \end{center}
\end{figure}
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.6]{figures/stream-analysis-viz-zoom-625.png}
      \caption[]{Zoomed in distance (JSD) signal and threshold for experiment 01}
      \label{fig:JSD-signal-zoom-01}
    \end{center}
\end{figure}
However, the change in JSD value is very abrupt as well. Almost instantly we go from close to 0 values to 1. We also had some false-positives, \textit{i.e.}, some JSD values above the threshold. This indicates to us that the window used is too small and sensitive to small changes, or that we can choose a more strict threshold.

@@

\subsubsection{Multiple distribution changes}
In this experiment, we wanted to see how our signal behaved when there were multiple change points, where each gaussian distribution was a scaled down version of the previous one. Figure \ref{fig:timeseries-r2} shows the time-series for the reference dataset R2 used in this experiment. It contained 5 million events with a single feature following a gaussian distribution with mean $\mu=100$ and standard deviation $\sigma=50$.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.6]{figures/timeseries-r2.png}
      \caption[]{Time-series for dataset R2}
      \label{fig:timeseries-r2}
    \end{center}
\end{figure}
The time-series for target dataset T2 is shown in Figure \ref{fig:timeseries-t2}. Notice that we change distributions at 2, 3 and 4 millionth events. Notice also that each subsequent distribution produces values whose domain is a subset of the previous distribution. 
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.6]{figures/timeseries-t2.png}
      \caption[]{Time-series for dataset T2}
      \label{fig:timeseries-t2}
    \end{center}
\end{figure}

@@

\subsubsection{Multiple distribution changes, each a superset of the previous one}
In this experiment, we wanted to do the reverse of the previous one. In this one, there are multiple change points but each new distribution spans a domain that is a superset of the previous one. Figure \ref{fig:timeseries-r3} shows the reference period used, with 5 million events and one feature following a gaussian distribution of mean $\mu=100$ and standard deviation $\sigma=10$.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.6]{figures/timeseries-r3.png}
      \caption[]{Time-series for dataset R3}
      \label{fig:timeseries-r3}
    \end{center}
\end{figure}
Figure \ref{fig:timeseries-t3} shows the target period used, also with 5 million events and one feature. Like in the previous experiment, we change distributions at the 2, 3 and 4 millionth event mark. On the other hand, unlike the previous experiment, we now make subsequent distributions cover the previous ones. We change the reference gaussian used with mean $\mu=100$ and standard deviation $\sigma=10$ at the 2, 3 and 4 millionth mark, to use standard deviations $\sigma=30$, $\sigma=40$ and $\sigma=50$, respectively.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.6]{figures/timeseries-t3.png}
      \caption[]{Time-series for dataset T3}
      \label{fig:timeseries-t3}
    \end{center}
\end{figure}

@@

\clearpage

@@

Figure \ref{fig:JSD-signal-test04} shows the computed JSD signal for the target period.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.6]{figures/stream-analysis-viz-test04.png}
      \caption[]{Distance (JSD) signal and threshold}
      \label{fig:JSD-signal-test04}
    \end{center}
\end{figure}
As expected, our JSD signal increases at the 2 millionth and 3 millionth event marks, because the distribution parameters change. At the 4 millionth event, we use the original reference distribution and notice that the JSD signal returns to values below the threshold, considering it a normal state, as we intended. Figure \ref{fig:JSD-signal-zoom-test04} is a zoom in on the first 2 millionth events, before any change occurs, and we see no false positives.

\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.6]{figures/stream-analysis-viz-zoom-test04.png}
      \caption[]{Zoomed in distance (JSD) signal and threshold}
      \label{fig:JSD-signal-zoom-test04}
    \end{center}
\end{figure}

@@

The first alert we obtain is:
\[1001000, [(0.0019999568566211323, x2)]\]
This tells us that at event 1,001,000 the only diverging feature in the system is feature \textit{x2} and what the p-value for the feature value at that time is about 0.2\%. This is consistent with our target period induced changes. We introduced a change in distribution at the 1 millionth event for feature x2. We caught this change 1000 events later because of our test frequency, which is adjustable to obtain results as recent as we need them to be.
